Prey movement shapes the development of predator expertise in a virtual bi-trophic system

### 2024-08-16 manuscript/manuscript.Rmd
updated discussion

Prey are known to use evasive tactics such as rapid escapes and movements to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], compelling predators to develop swift motor skills to succesfully capture such prey. However, predators may be limited in their capacity to develop the necessary attributes (e.g. physical, physiological, neurological) for fast-paced hunting, which may impair the development of their hunting expertise. There is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators. Experimental studies have shown that certain camouflage strategies in prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013]. For instance, @troscianko.etal2018 found that disruptive coloration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts. Given these examples, antipredator behaviour may also hold the potential to hinder the development of predator expertise.

@@

```{r}
# Import top 2 model fits
fit1 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-rank.rds"))
fit2 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-speed-rank.rds"))

@@

Predators usually optimize prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023]. Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise due to the challenges of investigating direct predator-prey interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience. Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the movement of the prey encountered.

Our results suggest that the acquisition of hunting expertise was honed through extensive practice. The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated. Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019]. However, the prey's speed was important in mediating this pattern at the population level. Encountering faster prey led to more difficult encounters for predators. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience. Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds. Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.

Prey speed also mediated differences among predator players in the development of their expertise. Individual predator players probably differed in their capacity to adjust to difficult prey. Animals are expected to have limited attention that constrain diet choice and image search formation [@dukas.etal2001]. Moreover, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011]. Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage. Parallel observations have been outlined in studies of prey camouflage strategies. For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation. Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies. Therefore, our observations suggest that prey antipredator behaviour can also impair predator expertise acquisition, with potential implications for predator-prey dynamics.

Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted, indicating the possible influence of other factors. It is hypothesized that longer time intervals between hunting events can hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions. Another potential reason why differences among individuals persisted is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival. For example, some players could experiment with the game out of boredom, which could impede ecologically realistic interpretations of our data. Lastly, the prey can be consumed in a match but carry that experience on to the next round without any penalty. Thus, interpretations of our results must be made with these caveats in mind. **inclure l'idée du prédateur qui peut aussi driver ça**

@@

We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise in a predator population. Our results suggest that predators gained expertise, as their success increased wit extensive practice, but that prey speed impaired expertise acquisition. Future analyses should investigate how efficient antipredator tactics are associated with prey experience, as this may reveal important insights on the eco-evolutionary dynamics of predator prey interactions. Lastly, virtual systems are increasingly used as systems to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

>>>>

Prey are known to use evasive tactics such as rapid escapes and movements to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], compelling predators to develop swift motor skills to succesfully capture such prey. However, predators may be limited in their capacity to develop the necessary attributes (e.g. physical, physiological, neurological) for fast-paced hunting, which may impair the development of their hunting expertise. There is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators. Experimental studies have shown that certain prey camouflage strategies can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013]. For instance, @troscianko.etal2018 found that disruptive coloration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts. Given these examples, antipredator behaviour may also hold the potential to hinder the development of predator expertise.

@@

```{r summary, cache = TRUE}
# Import top 2 model fits
fit1 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-rank.rds"))
fit2 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-speed-rank.rds"))

@@

Predators usually optimize prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023]. Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise due to the challenges of investigating direct predator-prey interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience. Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the speed of the prey encountered. Hence, our study provides empirical evidence that prey antipredator behaviour can impair the development of hunting expertise.

Our results suggest that the acquisition of hunting expertise was honed through extensive practice. The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated. These observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019]. However, the prey's speed was important in mediating this pattern at the population level. Encountering faster prey led to more difficult encounters for predators. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience. Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds. Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.

Prey speed also mediated differences among predator players in the development of their expertise, suggesting that individual predator players differed in their capacity to adjust to difficult prey. Animals are expected to have limited attention that constrain diet choice and image search formation, and hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@dukas.etal2001;@kelley.magurran2011]. Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage. Parallel observations have been outlined in studies of prey camouflage strategies. For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation. Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies. Therefore, our observations suggest that prey antipredator behaviour can also impair predator expertise acquisition, with potential implications for predator-prey dynamics.

Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted. A potential explanation is that the predators' hunting tactic may drive the antripredator tactic of the prey. We know from a previous study that predators specialize as cursorial or ambush hunters [@fraserfranco.etal2022]. Therefore, predators using a cursorial tactic may drive prey to move faster at the cost of having a slower acquisition of expertise.

Despite adjusting for the prey's speed, noticeable differences in the predators' expertise acquisition still emerged. One possible explanation is that the predators' hunting tactics may indirectly shape their own expertise through changes in prey behaviour. Previous research indicates that predators tend to specialize as either cursorial or ambush hunters in *DBD* [@fraserfranco.etal2022]. Consequently, predators employing a cursorial tactic may push prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey. An alternative explanation is that longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions. Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in *DBD*. As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival. For example, some players could experiment with the game out of boredom, which could shape how expertise is honed and impede ecologically realistic interpretations of our data.

@@

We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise acquisition in a human predator population in the game *Dead by Daylight*. As we did here with predators, future analyses should investigate how antipredator tactics are developped with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator prey interactions. Using a video game system, our study demonstrates that prey antipredator behavior can impair the acquisition of hunting expertise. This adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
